---
title: "Analysis of Tomasetti & Vogelstein (2015)"
knit: (function(inputFile, encoding) { input.dir <- normalizePath(dirname(inputFile)); for (section in list.files(pattern="analysis-.*?.Rmd")) { rmarkdown::render(section, encoding = "UTF-8", quiet=TRUE) }; rmarkdown::render(input = inputFile, encoding = encoding, quiet=TRUE, output_file = paste0(input.dir,'/../README.md')); rmarkdown::render(input = paste0(input.dir,'/../README.md'), encoding = encoding, output_format="html_document", output_file = paste0(input.dir,'/Tomasetti and Vogelstein analysis post.html')) })
output:
  md_document:
    variant: markdown_github
---


```{r echo=FALSE, results='asis', comment=''}
cat("#", rmarkdown::metadata$title)
```

To get an idea of the importance of what Tomasetti & Vogelstein (henceforth T&V) missed in their paper (and how important oversights may have been), I have compared the tissue types in recent CRUK cancer stats against those used in their paper, as the first of a few analyses. Source code (primarily within Rmarkdown documents), data *etc.* are available on GitHub [here](https://github.com/lmmx/tomasetti-vogelstein-2015).


```{r child = c("analysis-i.md", "analysis-ii.md", "analysis-iii.md")}
# The 'children' of the 'main' document can be specified (with the same result as the `include: after_body:` field in the YAML header) with either an explicit list:
#
# ```{r child = c("analysis-i.md", "analysis-ii.md", "analysis-iii.md")}
# ```
#
# or a call to `list.files`:
#
# ```{r child = list.files(pattern = 'analysis-.*?\\.md')}
# ```
# The former is preferable, but NB throws error if pattern doesn't match since child = character(0)
```
